---
title: "Final Project Part 2"
fontsize: 50
author: "Michelle Helfman"
date: "`r format(Sys.Date(),'%m-%d-%Y')`"
output:
  pdf_document: default
#bibliography: bibliography.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(lubridate)
library(scales)
library(pander)
library(tidyverse)
theme_set(theme_minimal())
options(scipen=999)
```

# Introduction  

Lotteries are everywhere.  You can buy tickets at grocery stores, gas stations, airports, kiosks at the mall, everywhere...  Lottery sales exceeded $210 billion in 2021, and individuals, families, friends, and groups of co-workers want their piece of the lottery pie.  I will be researching the two most popular lotteries in the United States:  Powerball and Mega Millions.  First, I will look into the most common numbers, jackpots, and the states where the winning tickets were sold for each lottery.  Then I will compare the Powerball and Mega Millions results to each other, focusing on commonalities and differences to answer questions about which lottery to participate in, which numbers to include, and which states to reside in when purchasing tickets.

Statistical analysis of the available lottery data gives insight into the patterns and frequency of the numbers drawn and the size and frequency of jackpots.  Then, the results can be displayed graphically for ease of understanding.


# Research Questions  
## Mega Millions Analysis
  1.  What are the 20 most frequent individual numbers?
  2.  What are the 20 largest jackpots? 
  3.  Which year had the paid out the most money since 2012?
  4.  How much time is between winning lottery drawings?
  5.  Which ten states have the most winners?
  
## Powerball Analysis
  1.  What are the 20 most frequent individual numbers?
  2.  What are the 20 largest jackpots? 
  3.  Which year had the paid out the most money since 2012?
  4.  How much time is between winning lottery drawings?
  5.  Which ten states have the most winners?
  
## Mega Millions vs. Powerball
  1. Compare the 10 most frequent individual numbers?
  2. Compare the winnings per year?
  3. Compare the 10 states have the most winners?
\newpage


# Addressing The Problem

My first step will be to extract lottery information from Mega Millions and Powerball 
related websites.  Then I will standardize the data for the 2 lotteries, including 
dates, numbers, and state information, then importing the results into R Markdown code 
to graphically display the analysis for the two lotteries.  Finally, I will compare 
the Mega Millions to Powerball results to see the commonalities and differences 
between the 2 lotteries and graphically depicting those results. 


# Approaching  The Problem

Once the data is standardized, it can be summarized, analyzed, and displayed for 
each lottery.  Having the information consistent across both lotteries, allows 
for comparison between the two lotteries, and the results are displayed similarly 
to the individual lotteries.


# Datasets (Select Description to Open Website)
The informataion was collected on October 15, 2022

***[Powerball drawings from Feburary 2010 to present](https://www.texaslottery.com/export/sites/lottery/Games/Powerball/Winning_Numbers/download.html)*** - This data has 11 columns including the drawing date that is broken into 3 
columns that will be concatenated into a usable date column.  In 2012, the Power 
Play was introduced so prior to that date, this is missing, but will not needed
for this project. Finally the resulting dataset will only include lottery 
drawings for January 2012 to present.

***[Mega Million drawings from December 2003 to present](https://www.texaslottery.com/export/sites/lottery/Games/Mega_Millions/Winning_Numbers/download.html)*** - This data has 11 columns including the drawing date that is broken 
into 3 columns that will be concatenated into a usable date column.  Finally the 
resulting dataset will only include lottery drawings for January 2012 to present.  

***[Powerball - Winning Jackpot and State information 2003 to present](https://www.powerball.net/winners)*** - This webpage has both state and 
jackpot information.  The dataset with state information has 3 columns.  The
jackpot related data has 6 columns.  Two columns have jackpot amounts, I will
be using the Annuity Value column since this is the publicized amount. 

***[Mega Millions - Winning Jackpot and State information from 2002 to present](https://mega-millions.com/winners)*** - This webpage has both state and 
jackpot information.  The dataset with state information has 3 columns.  The
jackpot related data has 6 columns.  The drawing date includes the day of
the week, which will need to be removed before standardizing this 
particular column.
	
\newpage
# Data Importing and Clensing
## Data Import
  1.  The lottery numbers for both the Mega Millions and Powerball drawings were 
  downloaded as CSV files from the Texas Lottery website.
  2.  Winning jackpot information for the Mega Millions lottery, including jackpot 
  amounts, dates of the winning drawings, and the states where the winners reside, 
  is copied from the official Mega Millions website.  The information is then pasted 
  into an excel spreadsheet, and a CSV file is created and additional commas in 
  the winner name were removed. 
  3.  Winning jackpot information for the Powerball lottery, including jackpot 
	amounts, dates of the winning drawings, and the states where the winners reside, 
	is copied from the official Powerball website.  The information is then pasted 
	into an excel spreadsheet, and a CSV file is created and additional commas in 
	the winner name were removed. 
  4.  The CSV files were loaded into a SQL Server database. 
  5.  Cleaned and standardized data will be loaded into R dataframes for selecting 
	%>% *specific columns* %>% filtering %>% grouping.

## Data Cleaning and Standardization
  1.  Stored Procedures were created to standardize the lottery drawing information 
  for both the Mega Millions and Powerball lotteries.  The drawing dates were 
  initially separated into the day, month, and year.  The 1st thing is to create 
  a combined date (mm/dd/yyyy), then use a cursor to create rows of just game_name, 
  drawing_date, and one number_drawn.  By adding a game name column, both lotteries 
  can be combined later into 1 dataset but still seen as two separate lotteries. 
  Finally, create a dataset for each lottery with information starting at 1/1/2012
  2.  the Mega Millions and Powerball state-related records required no additional 
	changes to the data, but a game name column was added so the records can be 
	combined later.
  3.  The Mega Millions jackpot records required removing the day of the week and 
	the abbreviations at the end of the day and converting this date to mm/dd/yyyy 
	format.  A game name column was added. 
  4.  The Powerbuilder jackpot records required no additional changes to the data, 
	but a game name column was added.
  5.  Lottery jackpot information was then added to the Mega Millions and Powerball 
	state records to create records with jackpot winner totals by state.


# Questions About Importing.
  1.  Can 2 files be loaded into 1 dataframe at the same time or does each one 
  have to be loaded into separate dataframes and combined using rbind()?
  2.  Can 2 sheets in an Excel spreadsheet be loaded into 1 dataframe at the same 
  time or does each one have to be loaded into separate dataframes and combined 
  using rbind()?

\newpage  
# Comments About Data Cleaning and Standardization
Start with the pieces that will become keys to tie the various tables or datasets 
together.  Ids (like customer numbers or codes) must match so the information is
meaningfully combined.  This may require the use of a cross-reference table.  
Then start standardizing dates, numbers (especially with decimal places), and any 
information used as filters, in groupings, or presented in reports.  Always keep 
in mind where the data is coming from and going to; indicators are needed to 
identify the system/dataset where the information belongs.


# Final Data Samples
```{r}
## Set the working directory
setwd("C:/Masters Degree/DSC 520 Statistics for Data Science/Final  Project")

## Mega Million Numbers Drawn
mm_numbers_df <- read.csv("Data Samples/Megamillions_Numbers_sample.csv")  
mm_numbers_df

## Powerball Numbers Drawn
pb_numbers_df <- read.csv("Data Samples/Powerball_Numbers_sample.csv")  
pb_numbers_df
```
\newpage  
```{r}
## Mega Million Winning Jackpots
mm_jackpot_df <- read.csv("Data Samples/Megamillions_Jackpot_sample.csv")  
mm_jackpot_df

## Powerball Winning Jackpots
pb_jackpot_df <- read.csv("Data Samples/Powerball_Jackpot_sample.csv")  
pb_jackpot_df

## Mega Million Jackpots by State
mm_state_df <- read.csv("Data Samples/Megamillions_States_with_Jackpots_sample.csv")
mm_state_df
```  
\newpage  
```{r}
## Powerball Jackpots by State
pb_state_df <- read.csv("Data Samples/Powerball_States_with_Jackpots_sample.csv")
pb_state_df
```  


# Uncovering New Information
By counting and summarizing the information will give insight into the patterns 
and frequency of the numbers drawn and the size and frequency of jackpots for each 
of the two lotteries.  Comparing and contrasting the 2 lotteries will show what 
number and winning states they have in common and which lottery has the largest 
jackpots.  


# Looking At The Data
This information will be grouped together by lottery (Mega Millions and Powerball).  
Individual lottery numbers will be counted.  Jackpots will be looked at individually, 
summed up by year and state.  The states where the winners reside are counted.  The
frequency between winning lotteries will averaged, the minimum and maximum time 
will be taken and normalized for graphing.  


# Slicing, Dicing, and Joining The Data
The information will be grouped by lottery (Mega Millions and Powerball).  The 
individual numbers (count), state jackpots totals (sum), lottery winners by state 
(top 20), and the largest jackpots (top 20)  The time between lotteries will be 
averaged, the minimun and maximum taken.  The 2 lotteries will be combined into 
dataframes by the most common drawn numbers, largest jackpots, and states with the
most winners to look at what is in common. New variables will be created for counts
and totals and new datafames will be created for records used in the graphs and 
tables used for analysis.


# Summarizing The Information
The information will be counted by the drawn numbers and states where the winners
reside.  Jackpots will be summarized by year and state and time between winners will
be normalized and plotted.  Taking the individual lottery results and then comparing 
the results to each other.  

\newpage
# Project Packages

  1. ggplot2 - For graphs
  2. plotly - For graphs
  3. stringr - String functions
  4. plyr - For manipulating data structures
  5. xlsx - Reading spreadsheets
  6. tidyverse - Variety of libraries for graphs, string manipulations, and more
  7. pander - Rendering R objects in RMarkdown
  8. lubridate - Working with dates
  9. Additional packages will be included, if needed.
  
# Plots and Tables
## Visualizations

  1. Bar Graphs (ggplot/geom_bar) - 
    i. Compare top 10 states with the most winners:  x axis - State, 
    y axis - Number of Winners, Legend - Lottery (Mega Millions and Powerball)
    ii. Compare top 10 most frequent numbers:  x axis - Number Drawn, 
    y axis - Number of Times Drawn, Legend - Lottery (Mega Millions and Powerball)
  2. Line Graphs (ggplot/geom_line) - 
    i. Total winnings by year for each lottery:  x axis - Year, y axis - Total Winnings
    ii Compare both lotteries total winnings per year: x axis - Year, 
    y axis - Total Winnings, Legend - Lottery (Mega Millions and Powerball)
  3. Histograms (ggplot/geom_histogram) - Time Between Winners per Lottery:  
  x axis - Time Between Winning Lotteries, y axis - Frequency
  
## Tables - by Lottery (Mega Millions and Powerball)

  1. 20 Most common Individual numbers - number drawn and count
  2. 20 Largest Jackpots - largest jackpots and drawing dates
  3. 10 States with the most winners state - states and number of winners
  
# Questions For Future Steps

  1. How do you depict frequency distribution?
  2. How can I incorporate machine learning?  Is there an appropriate location 
  where machine learning would be meaningful in predicting which lottery numbers
  to choose?
  
# Incorporate Machine Learning
After reading the upcoming assignments on Machine Learning, if there is an appropriate 
spot in predicting future lottery numbers I would like to incorporate this information.